manuscript

### 2025-01-31 doc/Manuscript.qmd
draft new intro

@@

We estimated survival probability, detection probability, recruitment of new frogs into the adult population, and adult population size for the Mossy Pond *R. sierrae* population using a site-specific Jolly-Seber CMR model implemented in a Bayesian framework with R statistical software [@R](v4.4.2). Our model was adapted from the mrmr R package [@joseph2019] (https://snarl1.github.io/mrmr/index.html), which was developed for the analysis of robust design CMR data collected from *R. sierrae* populations. The model represents a superpopulation of M individuals whose states (not recruited, alive, dead) and observations (detected, not detected) evolve through time, with any new frogs assumed to arise from local reproduction [@knapp2024]. We assessed model convergence via trace plots and Gelman–Rubin statistics (rhat).

>>>>

The well-documented decline of the endangered Sierra Nevada yellow-legegd frog in its namesake mountain range of California (bunch of citations) is emblematic of the crisis facing amphibians globally (citations). As is the case with numerous other species..the fungal pathogen Bd is the primary driver of SNYLF decline in the Sierra Nevada (citations), and the effects of climate change loom large, especially in an alpine environment expriencing accelerated climate change (mountain ranges are warming faster than other places?) (citation).

Though SNYLF decline fits the gloomy narrative, the examples of resilience and recovery that this species provides do not. Bd will typically wipe out (extirpate?) SNYLF populations (Vredenburg?), but a rare subset of individuals exhibit resistance to the pathogen ('resistant individuals') and contribute to the natural recovery of populations that persist in the presence of Bd (Knapp). These disease-persistent populations have become instrumental to conservation efforts for this species in their provision of resistant individuals as founders to establish new, resilient populations in previously-occupied habitat (Knapp 2024). We have the chance to pull off a rare feat: landscape scale recovery of an endangered species.

To successfully leverage these disease-persistent populations as a conservation resource, we must gain a detailed understanding of their demographics, dynamics, and the factors that influence their viability. Without good information, we can't make smart decisions and we will waste limited resources (Brown xxxx).

The population we focus on in this study-Mossy Pond-provides an excellent opportunity to fill the void of information surrounding this species in the northern portion of its range (cite interagency 2018 and Brown). Weird habitat, funky frogs, and little knowledge, etc..

In this investigation we set out to estimate demographic parameters (getting absolute abundance is really important, as Os for citations on that) and to investigate the rolethat climate plays in population viability. Ultimately, we hope our insights are useful for the management of this paprticualr population, for a deeper understanding of the species as a whole, and as a hint of how climate change may affect alpine amphibian populations in the future.

--------------------------------

By acting as donor populations for the human-facilitated reintroduction of disease-resistant animals into extirpated habitat, these unique Bd persistent populations have become instrumental to conservation efforts.

Some populations of MYLF exhibit resistence to Bd and will recover naturally (citations). These naturally-recovering populations (persistent populations) have become instrumental to human-facilitated conservation efforts, and present an opportunity for landscape scale recovery of an endangered species (Knapp 2024).

What effects may climate have on these recovery efforts?

-------------------------------------------

@@

We estimated survival probability, detection probability, recruitment of new frogs into the adult population, and adult population size for the Mossy Pond *R. sierrae* population using a site-specific Jolly-Seber CMR model implemented in a Bayesian framework with R statistical software [\@R](v4.4.2). Our model was adapted from the mrmr R package [@joseph2019] (https://snarl1.github.io/mrmr/index.html), which was developed for the analysis of robust design CMR data collected from *R. sierrae* populations. The model represents a superpopulation of M individuals whose states (not recruited, alive, dead) and observations (detected, not detected) evolve through time, with any new frogs assumed to arise from local reproduction [@knapp2024]. We assessed model convergence via trace plots and Gelman–Rubin statistics (rhat).
